On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2015-09-28 MS.tex
Comments of Emma and Mark added: - change terminology: juvenile ->seedling - check spelling Appendix E and D - change color: from green to blue for color-blind person

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 274/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate (GR) has often been seen as a central element of it's ecological strategy, and as ultimately linked to its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and GR, casting doubt on the capacity of traits to predict growth.

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate) using a meta-analysis of 112 studies ($>$ 500 correlations). To capture the effects of size, we tested for a shift in the direction of correlation between GR and traits across three plant stages: juvenile, sapling and adult.

\item Results were consistent with predictions. Specific leaf area was correlated with GR in juvenile but not in adult plants. Correlations of GR with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with GR only in one plant stage category: juveniles and adults, respectively.

@@

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass = SM) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') derived from a recently developed mechanistic model \citep{Falster:2011ii}?
\end{enumerate}
We gathered correlation coefficients between traits and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. juvenile, sapling and adult) as a surrogate for the effects of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.

@@

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``juvenile'', ``sapling'', or ``adult''.  We preferred the term ``juvenile'' to ``seedling'', because for a portion of the scientific community ``seedling'' also refers to a particular experimental condition. Using the above cut-offs we were also able to assign a stage category to studies for which no stage was distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 7 of 112 articles were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a particular stage category. Other equivocal cases were caused by ambiguity in the usage of ``adult'' by the authors. Some plants with a large stature were not described as adult by authors because they were still non reproductive \citep{King:2006dg,King:2006he}. Conversely plants with small stature -- because stunted by wind for example \citep{Stratton:2001ck}-- were sometimes assessed as adult when reproductive. For purposed of the present paper we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless of the plant's size or age. Results using the categories established by the authors are similar to those using the standardized categories (see Fig. \ref{fig:figA6} model.1). In the main text, we report only results using the standardized categories.

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a and d; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4} a-e). The correlation between growth and SLA shifted from strongly positive at juvenile stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation between growth and SM shifted from negative at juvenile stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult stage ($\bar{r}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{fig:fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages. The relationship between growth and SM in seedlings is expected to differ if absolute GR in mass is used instead of RGR mass. Here, for SM growth in mass was exclusively expressed in relative terms for seedlings (results shown in the Fig. \ref{fig:figA8} and Fig \ref{fig:figA4} f-j)

For WD, Hmax and Aarea, correlations with growth were also consistent with expectations, although with limitations arising from very unequal numbers of observations between growth stages for some traits. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$). At sapling and juvenile stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3} c), but also was consistent with no correlation (Fig \ref{fig:fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. We could not distinguish between these alternatives. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig. \ref{fig:fig3} b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

In total, we found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of $r$ reported for all traits. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}), and confounding with these factors is not a likely explanation fo the major results reported.

The results reported above were obtained using the "ideal" dataset, and were thereby confined to correlations across species, and under non-stressed conditions ($n = 176$). The ``complete'' dataset ($n = 216$) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This ``complete'' dataset generated similar effect sizes to those reported above from the ``ideal'' dataset, and led to similar conclusions (Fig. \ref{fig:fig3} grey points). Differences primarily concerned the confidence intervals around particular mean effect sizes in Fig. \ref{fig:fig3}. These differences in CI may have arisen from a variety of causes. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI 95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct}; these correlations unlike others used values of WD from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 8 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).

@@

As far as we are aware, our study is the first to compare Aarea and growth correlations between small and large plants. In the growth model (Appendix \ref{sec:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. Meta-analyses showed a positive correlation for adult and juvenile but not for saplings (n=2). In the juvenile stage, where most of data were available, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

Meta-analysis showed a positive correlation between Hmax and growth for adult stage, but not for saplings and juveniles. The trend for tall stature species to grow faster at a given tree size than small stature species has been noted across several tree size classes and forest types (e.g. \citealt{Poorter:2008iu,Wright:2010tp,Herault:2011dd,Ruger:2012jv,Iida:2014ep}). In \citet{Wright:2010tp}, Hmax was considered a weak predictor of growth rates of saplings; however it was the strongest predictor of growth rates of large trees. \citet{Poorter:2008iu} speculated that the Hmax-RGR relationship among large trees might arise from including small tree species with maximum size not much larger than the  10cm dbh lower threshold used to define large trees. In our growth equations, Hmax affects growth via the fraction of energy diverted to reproduction. At a given tree size, species closer to their Hmax are expected to be allocating relatively less to vegetative growth. However this applies only at sizes where some of the species have begun to reproduce, and should not affect sapling and juvenile growth rates.

@@

Clearly there is much scope for strengthening our understanding of correlations between traits and growth through (1) adding further empirical studies for traits and size-classes with weaker representation, (2) accounting more satisfactorily for variables such as light, and (3) expressing plant sizes on continuous scales of mass or height rather than through the approximate and partly subjective categories unavoidably adopted here. Additional data are specifically required in small plants (juvenile and sapling) for WD and Hmax, and in large plants (adult and sapling) for Aarea and more broadly for all traits requiring a measure of whole plant biomass (e.g. LMR, LAR, NAR). Nevertheless the existing published evidence, even with all its limitations, is consistent with predictions from a simple growth model, and this suggests that species traits can shed light on their growth trajectories, and hence on the changing growth rate rankings across species as individuals become larger.

@@

\clearpage
\begin{table}[h!]
\centering
\caption{Check-list of quality criteria for meta-analysis in ecology developed by Koricheva and Gurevitch (2014) }
{\footnotesize
\vspace{0.5cm}
\begin{tabular}{p{0.2cm}p{4cm}p{0.8cm}p{7cm}}
  \hline
  & Quality criteria & Check & How and where is it addressed? \\
  \hline
  1.& Reporting full details of bibliographic searches & yes	& We provided complete details of bibliographic searches (i.e. electronic reference data bases that were searched, keywords, language and search dates) in sub-section "Phase 2 - Systematic search" in material and method.\\
  2.& Reporting inclusion or exclusion criteria & yes & We specified several inclusion criteria in the sub-section "Phase 2 - Systematic search" in material and method. The Bibtex file `read.bib` on Guithub reports all the articles red but laid out because they failed to met our criteria.\\
  3.& Weighting effect sizes by study precision & yes & Effects sizes were weighted by number of species. See	 details in sub-section ``Effect size calculation'' in material and method.\\
 4.& Specifying meta-analytical model &	yes	& We used a mixed-effect model with a study ID as a random variation, and stage as a fixed effect. See details in sub-section "Stage effect" in material and methods. Results are shown in Fig. \ref{fig:fig3}.\\
 5.& Quantifying heterogeneity in effect sizes &	yes &	We estimated the extend of the heterogeneity using the statistics $I^{2}$ for each trait. See details of methods and results in Appendix E.\\
 6.& Exploring causes of heterogeneity & yes &	We explored the effect of stage, and the year of publication (see points 4 and 10 in this list). \\
 7.& Multifactorial analysis of explanatory variable	& yes & We checked if no one of several other explanatory moderators (i.e. growth measurement, experiment type or growth form) would not consistenly be a strongest predictor of the data. We did not test the multiple variables in a hierarchical fashion, because of the limited size of our dataset. See in Fig. \ref{fig:figA4} f-j and in Fig. \ref{fig:figA6}.\\
  8.& Testing for publication bias &	yes	& We used several methods (i.e. funnel plots, the ``trim and fill'' method and the Rosenberg’s fail-safe number). See details of methods and results in Appendix E and in Fig. \ref{fig:figA7}.\\
  9.& Sensitivity analysis &	yes &	See details in sub-section "Sensitivity analysis" in material and method. See results in Fig. \ref{fig:fig3} and Fig. \ref{fig:figA5}  \\
  10.& Exploring temporal changes in effect size &	yes	& See details of methods and results in Appendix E and Fig. \ref{fig:figA9}. \\
  11.& Controlling for phylogeny &	no &	It was not possible to control by the phylogeny since the number of species used to measured each coefficient of correlation ranged from 2 to 300 species.\\
  12.& Specifying the software used &	yes	& See sub-section ``Availability of code and data'' in material and methods\\
 13.& Providing reference list of primary studies included in the analysis &	yes &	See sub-section ``Availability of code and data'' in material and methods. The Bibtex file `meta-analyses.bib` corresponds to all the articles included in the meta-analyses. \\
  14.& Providing the data set used for meta-analysis  &	yes &	A copy of the complete data set is available on Github. See ``Availability of code and data'', section "Statistical analysis" in material and methods. \\
   \hline
\end{tabular}
}
\label{tab:meta}
\end{table}

@@

\textbf{Maximum height (Hmax):} Hmax affects growth via the fraction of energy allocated to vegetative growth versus to seed production ($dm_t/db$ in eq. \ref{eq:eq1}; Table \ref{tab:trade-offs}). For juveniles and seedlings this fraction is 1 for all species and so no effect is expected. But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates.

@@

\begin{table}[h!]
\centering
\caption{\textbf{List of the nineteen traits targeted}. Abbreviation,
definition and search term used in the meta-analysis}
\label{tab:lit_search}
\vspace{0.5cm}
{\footnotesize
\begin{tabular}{p{3cm}p{3cm}p{8cm}}
  \hline
Trait & Definition & Search terms \\
  \hline
Aarea & Photosynthetic rate per area & ``photosynthetic rate per area'' OR ``rate of CO2 assimilation per area'' OR ``Aarea'' OR ``Amax'' OR ``photosynthetic capacity per area'' \\
  Amass & Photosynthetic rate per mass & ``photosynthetic rate per mass'' OR ``rate of CO2 assimilation per mass'' OR ``Amass'' OR ``Amax'' \\
  Hmax & Maximum height & ``potential height'' OR ``maximum height'' OR ``Hmax'' \\
  Ks & Stem specific conductance & ``Ks'' OR ``stem specific conductance'' OR ``stem hydraulic conductance'' OR ``sapwood conductance'' \\
  LA & Leaf lamina area & ``leaf length'' OR ``leaf lamina area'' OR ``leaf width'' OR ``leaf size'' \\
  LAR & Leaf area ratio & ``leaf area ratio'' \\
  LMR & Leaf mass ratio & ``leaf mass ratio'' OR ``leaf mass fraction'' \\
  NARarea & Net assimilation rate per area &  ``net assimilation rate per area'' OR ``rate of dry mass increament per area'' OR ``NARarea'' OR ``NAR'' \\
  Narea & Leaf nitrogen content per area & ``Narea'' OR ``nitrogen per area'' OR ``N/area'' OR ``leaf nitrogen content per area'' OR ``LNCarea'' OR ``leaf nitrogen content" \\
  Nmass & Leaf nitrogen content per mass & ``Nmass'' OR ``nitrogen per mass'' OR ``N/mass'' OR ``leaf nitrogen content per mass'' OR ``LNCmass'' OR ``leaf nitrogen content" \\
  Parea & Leaf phosphorus content per area & ``Parea'' OR ``leaf P'' OR  ``leaf phosphorus content per area''  OR ``leaf phosphorus content'' \\
  Pmass & Leaf phosphorus content per mass & ``Pmass'' OR ``leaf P'' OR  ``leaf phosphorus content per mass'' OR ``leaf phosphorus content'' \\
  SA/LA & Sapwood area per leaf area & ``sapwood area per leaf area'' OR ``huber value'' OR ``leaf area per sapwood area'' \\
  Seed mass & Seed mass & ``seed mass'' OR ``seed size'' OR ``seed volume'' \\
  SLA & Specific Leaf Area & ``leaf mass per area'' OR ``specific leaf area'' OR ``leaf construction cost'' \\
  Thickness & Leaf thickness & ``leaf thickness''  OR ``leaf tissue density'' \\
  Vessel density & Vessel density & ``vessel density'' OR ``stem conduit density'' \\
  Vessel size & Vessel area or diameter & ``conduit area'' OR ``stem conduit area'' OR ``vessel diameter'' OR ``vessel size'' OR ``conduit size''   \\
  WD & Wood density & ``wood density'' OR``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'' OR ``SSD'' \\
   \hline
\end{tabular}
}
\end{table}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/FigA1.pdf}
\caption{\textbf{Allometric relationships between a) age and diameter (m), b) age and height (m), and c) diameter and height in trees according to the BAAD database.} Green and red lines are the cutoffs between stage categories used in the literature. Green lines corresponds to limits between juveniles and saplings, red lines to limits between saplings and adults.}
\label{fig:figA1}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/FigA2.pdf}
\caption{\textbf{Geographical distribution of site locations from studies performed in the field used in the meta-analysis}. Results shown here are from the ``complete'' dataset (see methods for details).}
\label{fig:figA2}
\end{figure}

>>>>

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 274/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate (GR) has often been seen as a central element of its ecological strategy, and as ultimately linked to its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and GR, casting doubt on the capacity of traits to predict growth.

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate) using a meta-analysis of 112 studies ($>$ 500 correlations). To capture the effects of size, we tested for a shift in the direction of correlation between GR and traits across three plant stages: seedling, sapling and adult.

\item Results were consistent with predictions. Specific leaf area was correlated with GR in seedling but not in adult plants. Correlations of GR with wood density and assimilation rate were not affected by stage. Seed mass and maximum height were correlated with GR only in one plant stage category: seedlings and adults, respectively.

@@

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass = SM) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') derived from a recently developed mechanistic model \citep{Falster:2011ii}?
\end{enumerate}
We gathered correlation coefficients between traits and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. seedling, sapling and adult) as a surrogate for the effects of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.

@@

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``seedling'', ``sapling'', or ``adult''. Using the above cut-offs we were also able to assign a stage category to studies for which no stage was distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 7 of 112 articles were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a particular stage category. Other equivocal cases were caused by ambiguity in the usage of ``adult'' by the authors. Some plants with a large stature were not described as adult by authors because they were still non reproductive \citep{King:2006dg,King:2006he}. Conversely plants with small stature -- because stunted by wind for example \citep{Stratton:2001ck}-- were sometimes assessed as adult when reproductive. For purposed of the present paper we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless of the plant's size or age. Results using the categories established by the authors are similar to those using the standardized categories (see Fig. \ref{fig:figA6} model.1). In the main text, we report only results using the standardized categories.

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a and d; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4} a-e). The correlation between growth and SLA shifted from strongly positive at seedling stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation between growth and SM shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult stage ($\bar{r}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{fig:fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between seedling and adult stages. The relationship between growth and SM in seedlings is expected to differ if absolute GR in mass is used instead of RGR mass. Here, for SM growth in mass was exclusively expressed in relative terms for seedlings (results shown in the Fig. \ref{fig:figA8} and Fig \ref{fig:figA4} f-j)

For WD, Hmax and Aarea, correlations with growth were also consistent with expectations, although with limitations arising from very unequal numbers of observations between growth stages for some traits. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$). At sapling and seedling stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3} c), but also was consistent with no correlation (Fig \ref{fig:fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. We could not distinguish between these alternatives. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig. \ref{fig:fig3} b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

In total, we found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of $r$ reported for all traits. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}), and confounding with these factors is not a likely explanation of the major results reported.

The results reported above were obtained using the "ideal" dataset, and were thereby confined to correlations across species, and under non-stressed conditions ($n = 176$). The ``complete'' dataset ($n = 216$) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This ``complete'' dataset generated similar effect sizes to those reported above from the ``ideal'' dataset, and led to similar conclusions (Fig. \ref{fig:fig3} grey points). Differences primarily concerned the confidence intervals around particular mean effect sizes in Fig. \ref{fig:fig3}. These differences in CI may have arisen from a variety of causes. For instance, the correlation between growth and WD at seedling stage shifted from negative to non significant (i.e CI 95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct}; these correlations unlike others used values of WD from adult plants together with growth from seedlings. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 8 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).

@@

As far as we are aware, our study is the first to compare Aarea and growth correlations between small and large plants. In the growth model (Appendix \ref{sec:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. Meta-analyses showed a positive correlation for adult and seedling but not for saplings (n=2). In the seedling stage, where most of data were available, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

Meta-analysis showed a positive correlation between Hmax and growth for adult stage, but not for saplings and seedlings. The trend for tall stature species to grow faster at a given tree size than small stature species has been noted across several tree size classes and forest types (e.g. \citealt{Thomas:1996do,Poorter:2008iu,Wright:2010tp,Herault:2011dd,Ruger:2012jv,Iida:2014ep}). In \citet{Wright:2010tp}, Hmax was considered a weak predictor of growth rates of saplings; however it was the strongest predictor of growth rates of large trees. \citet{Poorter:2008iu} speculated that the Hmax-RGR relationship among large trees might arise from including small tree species with maximum size not much larger than the 10cm dbh lower threshold used to define large trees. In our growth equations, Hmax affects growth via the fraction of energy diverted to reproduction. At a given tree size, species closer to their Hmax are expected to be allocating relatively less to vegetative growth. However this applies only at sizes where some of the species have begun to reproduce, and should not affect sapling and seedling growth rates.

@@

Clearly there is much scope for strengthening our understanding of correlations between traits and growth through (1) adding further empirical studies for traits and size-classes with weaker representation, (2) accounting more satisfactorily for variables such as light, and (3) expressing plant sizes on continuous scales of mass or height rather than through the approximate and partly subjective categories unavoidably adopted here. Additional data are specifically required in small plants (seedling and sapling) for WD and Hmax, and in large plants (adult and sapling) for Aarea and more broadly for all traits requiring a measure of whole plant biomass (e.g. LMR, LAR, NAR). Nevertheless the existing published evidence, even with all its limitations, is consistent with predictions from a simple growth model, and this suggests that species traits can shed light on their growth trajectories, and hence on the changing growth rate rankings across species as individuals become larger.

@@

\clearpage
\begin{table}[h!]
\centering
\caption{Check-list of quality criteria for meta-analysis in ecology developed by Koricheva and Gurevitch (2014) }
{\footnotesize
\vspace{0.5cm}
\begin{tabular}{p{0.2cm}p{4cm}p{0.8cm}p{7cm}}
  \hline
  & Quality criteria & Check & How and where is it addressed? \\
  \hline
  1.& Reporting full details of bibliographic searches & yes	& We provided complete details of bibliographic searches (i.e. electronic reference data bases that were searched, keywords, language and search dates) in sub-section "Phase 2 - Systematic search" in material and method.\\
  2.& Reporting inclusion or exclusion criteria & yes & We specified several inclusion criteria in the sub-section "Phase 2 - Systematic search" in material and method. The Bibtex file `read.bib` on Github reports all the articles red but laid out because they failed to met our criteria.\\
  3.& Weighting effect sizes by study precision & yes & Effects sizes were weighted by number of species. See	 details in sub-section ``Effect size calculation'' in material and method.\\
 4.& Specifying meta-analytical model &	yes	& We used a mixed-effect model with a study ID as a random variation, and stage as a fixed effect. See details in sub-section "Stage effect" in material and methods. Results are shown in Fig. \ref{fig:fig3}.\\
 5.& Quantifying heterogeneity in effect sizes &	yes &	We estimated the extend of the heterogeneity using the statistics $I^{2}$ for each trait. See details of methods and results in Appendix E.\\
 6.& Exploring causes of heterogeneity & yes &	We explored the effect of stage, and the year of publication (see points 4 and 10 in this list). \\
 7.& Multifactorial analysis of explanatory variable	& yes & We checked if no one of several other explanatory moderators (i.e. growth measurement, experiment type or growth form) would not consistenly be a strongest predictor of the data. We did not test the multiple variables in a hierarchical fashion, because of the limited size of our dataset. See in Fig. \ref{fig:figA4} f-j and in Fig. \ref{fig:figA6}.\\
  8.& Testing for publication bias &	yes	& We used several methods (i.e. funnel plots, the ``trim and fill'' method and the Rosenberg’s fail-safe number). See details of methods and results in Appendix E and in Fig. \ref{fig:figA7}.\\
  9.& Sensitivity analysis &	yes &	See details in sub-section "Sensitivity analysis" in material and method. See results in Fig. \ref{fig:fig3} and Fig. \ref{fig:figA5}  \\
  10.& Exploring temporal changes in effect size &	yes	& See details of methods and results in Appendix E and Fig. \ref{fig:figA9}. \\
  11.& Controlling for phylogeny &	no &	It was not possible to control by the phylogeny since the number of species used to measured each coefficient of correlation ranged from 2 to 300 species.\\
  12.& Specifying the software used &	yes	& See sub-section ``Availability of code and data'' in material and methods\\
 13.& Providing reference list of primary studies included in the analysis &	yes &	See sub-section ``Availability of code and data'' in material and methods. The Bibtex file `meta-analyses.bib` corresponds to all the articles included in the meta-analyses. \\
  14.& Providing the data set used for meta-analysis  &	yes &	A copy of the complete data set is available on Github. See ``Availability of code and data'', section "Statistical analysis" in material and methods. \\
   \hline
\end{tabular}
}
\label{tab:meta}
\end{table}

@@

\textbf{Maximum height (Hmax):} Hmax affects growth via the fraction of energy allocated to vegetative growth versus to seed production ($dm_t/db$ in eq. \ref{eq:eq1}; Table \ref{tab:trade-offs}). For seedlings and seedlings this fraction is 1 for all species and so no effect is expected. But once some species begin to produce seed, then species closer to their Hmax value are expected to be allocating more to seed production and less to vegetative growth. Hence when comparing across plants at a given size, species with smaller Hmax are expected to show slower vegetative growth rates.

@@

\begin{table}[h!]
\centering
\caption{List of the nineteen traits targeted. Abbreviation,
definition and search term used in the meta-analysis}
\label{tab:lit_search}
\vspace{0.5cm}
{\footnotesize
\begin{tabular}{p{3cm}p{3cm}p{8cm}}
  \hline
Trait & Definition & Search terms \\
  \hline
Aarea & Photosynthetic rate per area & ``photosynthetic rate per area'' OR ``rate of CO2 assimilation per area'' OR ``Aarea'' OR ``Amax'' OR ``photosynthetic capacity per area'' \\
  Amass & Photosynthetic rate per mass & ``photosynthetic rate per mass'' OR ``rate of CO2 assimilation per mass'' OR ``Amass'' OR ``Amax'' \\
  Hmax & Maximum height & ``potential height'' OR ``maximum height'' OR ``Hmax'' \\
  Ks & Stem specific conductance & ``Ks'' OR ``stem specific conductance'' OR ``stem hydraulic conductance'' OR ``sapwood conductance'' \\
  LA & Leaf lamina area & ``leaf length'' OR ``leaf lamina area'' OR ``leaf width'' OR ``leaf size'' \\
  LAR & Leaf area ratio & ``leaf area ratio'' \\
  LMR & Leaf mass ratio & ``leaf mass ratio'' OR ``leaf mass fraction'' \\
  NARarea & Net assimilation rate per area &  ``net assimilation rate per area'' OR ``rate of dry mass increament per area'' OR ``NARarea'' OR ``NAR'' \\
  Narea & Leaf nitrogen content per area & ``Narea'' OR ``nitrogen per area'' OR ``N/area'' OR ``leaf nitrogen content per area'' OR ``LNCarea'' OR ``leaf nitrogen content" \\
  Nmass & Leaf nitrogen content per mass & ``Nmass'' OR ``nitrogen per mass'' OR ``N/mass'' OR ``leaf nitrogen content per mass'' OR ``LNCmass'' OR ``leaf nitrogen content" \\
  Parea & Leaf phosphorus content per area & ``Parea'' OR ``leaf P'' OR  ``leaf phosphorus content per area''  OR ``leaf phosphorus content'' \\
  Pmass & Leaf phosphorus content per mass & ``Pmass'' OR ``leaf P'' OR  ``leaf phosphorus content per mass'' OR ``leaf phosphorus content'' \\
  SA/LA & Sapwood area per leaf area & ``sapwood area per leaf area'' OR ``huber value'' OR ``leaf area per sapwood area'' \\
  Seed mass & Seed mass & ``seed mass'' OR ``seed size'' OR ``seed volume'' \\
  SLA & Specific Leaf Area & ``leaf mass per area'' OR ``specific leaf area'' OR ``leaf construction cost'' \\
  Thickness & Leaf thickness & ``leaf thickness''  OR ``leaf tissue density'' \\
  Vessel density & Vessel density & ``vessel density'' OR ``stem conduit density'' \\
  Vessel size & Vessel area or diameter & ``conduit area'' OR ``stem conduit area'' OR ``vessel diameter'' OR ``vessel size'' OR ``conduit size''   \\
  WD & Wood density & ``wood density'' OR``WD'' OR ``wood specific gravity'' OR ``stem density'' OR ``stem specific density'' OR ``SSD'' \\
   \hline
\end{tabular}
}
\end{table}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/FigA1.pdf}
\caption{Allometric relationships between a) age and diameter (m), b) age and height (m), and c) diameter and height in trees according to the BAAD database. Blue and red lines are the cut-offs between stage categories used in the literature. Blue lines corresponds to limits between seedlings and saplings, red lines to limits between saplings and adults.}
\label{fig:figA1}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/FigA2.pdf}
\caption{Geographical distribution of site locations from studies performed in the field used in the meta-analysis. Results shown here are from the ``complete'' dataset (see methods for details).}
\label{fig:figA2}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/FigA3.pdf}
\caption{Range of correlations (r) between
growth and traits reported in the literature. Plots show the coefficients of correlation (r)
between plant growth and a) Specific leaf area (SLA, n =
111), b) Wood density (WD, n = 64), c) Asymptotic
maximum height (Hmax, n = 23), d) Seed mass (Seedmass, n = 36 and
e) Rate of CO2 assimilation (Aarea, n = 24) reported from 112
articles. The x axis represents the article ID sorted by the average reported coefficient value r. Each bubble represents an r value, the diameter of the bubble is proportional to its weight in the meta-analysis (i.e. number of species). The colours correspond to the stage for which the r values have been recorded: red points correspond to adult stage, orange points to saplings, and blue points to seedlings. Results shown here are from the ``complete'' dataset (see methods for details). }
\label{fig:figA3}
\end{figure}
